---
title: "A simple web app with shiny"
author: "Marie-Louise Timcke"
date: "28. März 2016"
output: html_document
---
This is an R Markdown file containing the code for this web app on Germany's air pollutants emissions. If you're new to R: Don't panic! This may look complicated at first (not only because the code is squeezed to be a little bit more readable in the small space we have for the iframe on our website), but all it takes to understand this are our tutorials regarding:
[shiny](http://journocode.com/2016/03/27/r-your-first-web-application-with-shiny/), [ggplot2](http://journocode.com/2016/03/02/r-the-ggplot2-package/) and [tidy data](http://journocode.com/2016/03/05/r-tidy-data/)!

Visit us on [GitHub](https://github.com/journocode/shinyApp) to find a markdown, the data and the raw code of this app.

**Code**

Load required packages
```{r}
library(shiny) 
library(readxl) # to load the data into R
library(tidyr) # to restructure the data
library(ggplot2) # for the barchart
library(dplyr) # to group and filter the data
```

Read the Excel-file and melt it. The vector *pollutantsorder* will help us to sort the select options in the app the way we want.
Don't forget to set the working directory correctly!
```{r}
emissions <- read_excel("TotalEmissionsGermanyInKT.xlsx")

emissions_restructed <- emissions %>% 
  tidyr::gather(key = "year", value = "emissions", 2:10)

pollutantsorder <- c("TSP", "SO2", "PM 2.5", "PM 10", 
                     "NOx", "NMVOCs", "NH3", "CO")
```

Build the user interface
```{r}
ui <- shinyUI(
 fluidPage(
    # header panel
 fluidRow(column(width = 10, style = "font-size: 25pt; 
                                     line-height: 40pt; 
                                     width = 100",
              tags$strong("Germanys air 
                          pollutants emissions")), 
       column(width = 2,
             tags$head(tags$img(src='http://journocode.
                                com/wordpress/wp-content
                                /uploads/2016/
                                01/small-logo.png', 
                                align = "left", 
                                width= "100")))
          ),
# sidebar with user input options
sidebarPanel(style = "background-color: #78d9c9;",
              tags$style(type='text/css', 
                         ".selectize-input { 
                         font-size: 12pt; 
                         line-height: 13pt;} 
                         .selectize-dropdown {
                         font-size: 12pt; 
                         line-height: 13pt;}"),
              width = 3,
              # checkbox for pollutants
              checkboxGroupInput("pollutants",
                                 label = HTML(
                                 '<p style="color:
                                 white; 
                                 font-size:12pt">
                                 Pollutants</p>'), 
              choices = pollutantsorder, 
              selected = pollutantsorder),
              # select option for years
              selectInput("year",
                         label = HTML(
                         '<p style="color:white; 
                         font-size: 12pt"> Year </p>'), 
                         choices = unique(
                         emissions_restructed$year), 
                         multiple = TRUE, 
                         selected = c(2005:2013)),
              # text with additional information
              helpText(
                       HTML('<p style="color:white; 
                           font-size: 9pt">choose air 
                           pollutants by clicking 
                           the check boxes,
                           exclude years with a 
                           click and the 
                           backspace key</p>'))
              ),
# main panel with output in different tabs
mainPanel(tabsetPanel(type = "tabs", 
                     # tab with plot output
                      tabPanel("Graphic",
                               plotOutput("mplot"), 
                               style = "width:100%"), 
                      # tab with data output
                      tabPanel("Data",
                               htmlOutput("text"), 
                               tableOutput("table"), 
                               style = "font-size:70%",
                               htmlOutput(
                               "datasource")),
                      # tab with this R Markdown file
                      tabPanel("Code",
                      includeMarkdown(
                      "emissions_app.Rmd"))
    )
    )
  ))
```

Build the server part
```{r}
server <- shinyServer(function(input, output) {
data <- reactive({ # reactive part = thos code is repeated when user input changes
    
 validate( # define error messages if user doesn't choose anything
  need(input$pollutants != "", 
        "Please select at least one air polutant"),
  need(input$year != "", 
       "Please select at least one year")
)
    
  # filter data according to user input
  plotdata <- emissions_restructed  %>%
    as.data.frame() %>%
    filter(
      pollutants %in% input$pollutants 
      & year %in% input$year)
   
  # define reactive scale for y-axis
  scalecalc <- plotdata %>%
    group_by(year) %>%
    summarize(value = sum(emissions))
    
  scalemax <- max(scalecalc$value)
  scalesteps <- round(scalemax/5, digits = -1)
    
  list(plotdata = plotdata,
       scalemax = scalemax,
       scalesteps = scalesteps
       )
})
# build plot output with ggplot2
output$mplot <- renderPlot({
  myplot <- ggplot(data = data()$plotdata, 
                   aes(as.factor(year), y = emissions, 
                       fill = factor(pollutants), 
                       order = pollutants)) + 
  geom_bar(stat = "identity") + 
  xlab("Year") + 
  ylab("Emissions in kt") +
  theme_minimal()+
  ggtitle("Barchart of air pollutants emissions\n") +
  guides(fill=guide_legend(title="pollutants", 
                            reverse = T))+
  scale_y_continuous(breaks = 
                     seq(0,data()$scalemax,   
                         data()$scalesteps),
                     labels = abs(seq(0,
                                  data()$scalemax,
                                  data()$scalesteps))) +
  theme(plot.title = 
          element_text(family="Arial", 
                       face="bold", 
                       size=18),
        axis.text.x = 
          element_text(angle = 0, 
                       family="Arial", 
                       size=13),                 
        axis.text.y = 
          element_text(angle = 0, 
                       family="Arial", 
                       size=13),
        axis.title.x = 
          element_text(size=14, 
                       face="bold", 
                       vjust = -1),
        axis.title.y = 
          element_text(size=14, 
                       face="bold", 
                       vjust = 2)
          ) +
  scale_fill_manual(values = c("TSP" = "#ffc7e4", 
                                "SO2" = "#ffb155", 
                                "PM 2.5" = "#ff6f69", 
                                "PM 10" = "#b1e6e6", 
                                "NOx" = "#77b1d5", 
                                "NMVOCs" = "#c0b7db", 
                                "NH3" = "#fcffaf", 
                                "CO" = "#78d9c9"), 
                                drop = F)
    
print(myplot)
})
# build table output
output$table <- renderTable({
  emissions_restructed %>%
  filter(pollutants %in% input$pollutants & 
         year %in% input$year) %>%
  spread(year, emissions) # spread() = opposite of gather() 
})
# build text output for table source
output$datasource <- renderUI({
  tags$div(
     tags$strong("Source:"), 
     tags$a("Umweltbundesamt", 
            href="http://www.umweltbundesamt.de/
                  daten/luftbelastung/
                  luftschadstoff-emissionen-
                  in-deutschland")
  )       
})  
# text output for data description
output$text <- renderUI({
  tags$div(
    HTML('<p style="color:black; font-size: 9pt">
         This data on Germanys air pollutants 
         emissions was downloaded from the german
         Federal Environment Agencys website. The 
         table provides information on nitrogen oxides 
         (NOx), ammonia (NH3), volatile organic compounds            
         without methane (NMVOC), sulfur dioxide (SO2) 
         and dust - including the fine dust fractions 
         PM 10 and PM 2.5 - and carbon monoxide (CO).</p>')
    )   
  })
})
```

Run the app
```{r}
shinyApp(ui = ui, server = server)
```


